the choice of softImpte lambda

### 2018-01-27 analysis/softImputelambda.Rmd
Files commited by wflow_commit.

---
title: "apply flashr2 with OCV"
author: "wei wang"
date: 09-30
output: 
  html_document:
    code_folding: hide
---

@@

## preparation

1. creat a folder

```
mkdir Breastcancer
```

2. put your data into this folder

here we use the file 'example.mat'. the reason we use `.mat` file is because there is a matlab package in our comparision.

3. creat the wrapper function in R and matlab

### wrapper function for OCV

@@

PMA.wrapper = function(Y,ngrids = 10,K,fold = 10){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.3,0.8,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
  res_log = capture.output({out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)})
  return(list(d = out$d, u = out$u, v = out$v))
}

@@

NSF_OCV = function(OCVindex){
  if(file.exists("NBSFAout")){
    unlink("NBSFAout", recursive= T)
  }
  system("mkdir NBSFAout")
  # write the missing index

  writeMat("./NBSFAout/OCVindex.mat", OCVindex = OCVindex)

  system("matlab -nosplash -nodesktop -r \"addpath(\'../\'); run(\'missvalue.m\');exit;\" ")
  g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
  nsfa_rmse = as.vector(g_nsfa$RMSE)
  return(nsfa_rmse)
}

@@

This file name is `wrapper.R`

### matlab code

In this folder

```{}
ddpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');

% you need to change the path and the data file (not centered)
load ../Ydata.mat;
Ycentered=Y;
settings=defaultsettings();
[settings.D,settings.N]=size(Ycentered);
settings.iterations=100;
% now the missing is decide by the this is the missing index.(or you can upload the missing matrix)

file_path = fullfile(pwd,'NBSFAout/OCVindex.mat');
load(file_path);

% missindex = ones(settings.D,settings.N); this is not missing index this
% is observed index.
missindex = (~isnan(Y));
k_fold = 10;
foldindex = zeros(k_fold,k_fold,2);
for k = 1:2
    for j = 1:k_fold
        for i = 1:k_fold
            foldindex(i,j,k) = OCVindex.foldindex(i + k_fold*(j - 1) + k_fold * k_fold * (k - 1));
        end
    end
end

SSE = zeros(1,k_fold);
for i = 1:k_fold
    mvmask = missindex;
    for j = 1:k_fold
        mvmask(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2))) = 0;
    end
    mvmask(:,sum(mvmask) == 0) = (missindex(:,sum(mvmask) == 0));
    initialsample=init_nsfa(settings);
    [finalsample,resultstable]=nsfa(Ycentered,mvmask,initialsample,settings);
    Y_hat = (finalsample.G *  finalsample.X);
    for j = 1:k_fold
        SSE(i) = SSE(i) + ...
            sum(nansum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
            Y_hat(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))).^2));
    end

end

RMSE = sqrt(sum(SSE)/(settings.D*settings.N));
RMSE;

save_path = fullfile(pwd,'NBSFAout/NSFAresult');
save(save_path,'RMSE');
```

you need change the path `~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/` and `~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/` which have the NBSFA matlab code.

### creat the run.R file to run the code

In this case, we have problem in running `flashr2` on RCC. so we just try other methods first.

```{}
source("../wrapper.R")
library(R.matlab)
## run the code
Y_centered = readMat("../example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
OCVindex = OCV_index(Y,k_fold = 10)
OCVdata = OCV_data(Y,OCVindex,k_fold = 10)

PN_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "PN",Kmax = 40)
flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash",Kmax = 40)
flashGwn_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash_wn",Kmax = 40)
flashB_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash_bf",Kmax = 40)
pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "PMD",Kmax = 40)
soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "softImpute",Kmax = 40)
nsfa_rmse = NSF_OCV(OCVindex)

result = c(PN_rmse,flashG_rmse,flashGwn_rmse,flashB_rmse,nsfa_rmse,pmd_rmse,soft_rmse)
saveRDS(result, "./NBSFAout/output.rds")
print(sessionInfo())
```

this file name is `runOCV.R`

### creat folder to track the result

```
mkdir outlog
```

### creat .sbatch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=./outlog/arrayJob_%A_%a.out
#SBATCH --error=./outlog/arrayJob_%A_%a.err
#SBATCH --array=1-20
#SBATCH --time=26:00:00
#SBATCH --partition=mstephens
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=8000

######################
# Begin work section #
######################

# Print this sub-job's task ID
module load R
module load matlab

mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose ../runOCV.R
```

this file is name as `OCV.sbatch`

## run OCV

now in your folder, you should have:

```
[weidong@midway2-login1 Breastcancer]$ ls
OCV.sbatch  example.mat  missvalue.m  outlog  runOCV.R  wrapper.R
```

```
sbatch OCV.sbatch
```

## to get the result

```{}
T = 20
results = matrix(NA,ncol = 7, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_folder = "NBSFAout"
  out_file = "output.rds"
  file_name = file.path(test_folder,out_folder,out_file)
  results[i,] = try(readRDS(file_name))
}

saveRDS(results,"./boxplot.rds")
```

## plot the result

```{r}
# this is the independent result

library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none", x_label){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(x_label, each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
    geom_boxplot()+
    # geom_violin()+
    ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15),
          plot.title = element_text(size = 15, face = "bold"),
          axis.text.y = element_text(size =12.6),
          axis.text.x = element_text(size =12.6,angle = 45, hjust = 1))
  p
}

setwd("~/HG/flash_workflow/analysis")

##############  MovieLens data

ML_res = readRDS("../data/output/missingdata/MovieLens/ML100K_box.rds")
ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
ML_res = matrix(as.numeric(ML_res),ncol = 4)
# c("flash","NBSFA","PMD","softImpute")
ml_res = readRDS("../data/output/missingdata/MovieLens/boxplot_cv.rds")
# flashG_rmse,flashGwn_rmse,pmd_rmse,soft_rmse
ml_res[1:20,] = sqrt((ml_res[1:20,])^2 * (943*1682)/(100000))
ml_res = ml_res[1:20,]
ML_res = ML_res[-c(2,13,17,21,29,37,62,76,77,93,95,100),]
ML_res = ML_res[1:20,]
# add PN result
ML_pn_res = readRDS("../data/output/missingdata/MovieLens/box_cv_pn.rds")
# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(ML_pn_res[,1],ML_pn_res[,2],ML_res[,3],ml_res[,3],ML_res[,2],ML_res[,4],ml_res[,4])

x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotM = plot_res(output,"MovieLens data",x_label = x_label)

############ Breast Cancer data

BC_res = readRDS("../data/output/missingdata/BreastCancer/BreastCancer_box.rds")
BC_res = BC_res[1:20,]
# c("flash","NBSFA","PMD","softImpute")
bc_res = readRDS("../data/output/missingdata/BreastCancer/box_Breast.rds")
# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(bc_res[,2],bc_res[,1],BC_res[,3],bc_res[,6],BC_res[,2],BC_res[,4],bc_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotB = plot_res(output,"Breast Cancer data",x_label = x_label)

############ GTEx data

GZ_res = readRDS("../data/output/missingdata/GTExZsocre/gtexzscore_box.rds")
GZ_res = GZ_res[1:20,]
# c("flash","NBSFA","PMD","softImpute")
gz_res = readRDS("../data/output/missingdata/GTExZsocre/box_cv.rds")
# flashG_rmse,nsfa_rmse,pmd_rmse,soft_rmse
gz_pn_res = readRDS("../data/output/missingdata/GTExZsocre/box_pn.rds")
gz_pn_res[c(23,25,29,34),] = NA
gz_pn_res = matrix(as.numeric(gz_pn_res[21:40,]),ncol = 2)

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(gz_res[,1],gz_pn_res[,2],GZ_res[,3],gz_res[,3],GZ_res[,2],GZ_res[,4],gz_res[,4])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotG = plot_res(output,"GTEx data",x_label = x_label)

############ Text data

PT_res = readRDS("../data/output/missingdata/DenoiseRtext/president_box.rds")
PT_res = PT_res[1:20,]
# c("flash","NBSFA","PMD","softImpute")
pt_res = readRDS("../data/output/missingdata/DenoiseRtext/box_president.rds")
# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(pt_res[,3],pt_res[,1],PT_res[,3],pt_res[,6],PT_res[,2],PT_res[,4],pt_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotP = plot_res(output,"Text data",x_label = x_label)

############ Tumor data

DT_res = readRDS("../data/output/missingdata/DenoiseRtumor/denoiseRtumor_box.rds")
DT_res = DT_res[1:20,]
# c("flash","NBSFA","PMD","softImpute")
dt_res = readRDS("../data/output/missingdata/DenoiseRtumor/box_denoiseTumor.rds")
# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(dt_res[,2],dt_res[,1],DT_res[,3],dt_res[,6],DT_res[,2],DT_res[,4],dt_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotT = plot_res(output,"Tumor data",x_label = x_label)

# gridExtra::grid.arrange(plotP,plotT,plotG,plotB,plotM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
gridExtra::grid.arrange(plotM,plotT,plotG,plotB,plotP, layout_matrix = rbind(c(1,2,3),c(4,5,NA) ))

## other information

please click here to check other information [Missing data OCV settings](OCVmissdataSetting.html)

>>>>

---
title: "the choice of softImpte lambda"
author: "Wei Wang"
date: 2018-01-27
output: 
  html_document:
    code_folding: hide
---

@@

## Document

Based on the document of the authors ![https://cran.r-project.org/web/packages/softImpute/softImpute.pdf], we should choose lambda ($\lambda$) as "Ideally lambda should be chosen so that the solution reached has rank slightly
less than rank.max."

## Brease Cancer data in flash paper

@@

@@

@@

read the data

```{r}
library(R.matlab)
## run the code
Y_centered = readMat("../data/output/missingdata/BreastCancer/example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
```

first we try to get the lambda based on the authors suggestion.

```{r}
OCVindex = OCV_index(Y,k_fold = 10)
OCVdata = OCV_data(Y,OCVindex,k_fold = 10)

colindex = OCVindex$colindex
rowindex = OCVindex$rowindex
foldindex = OCVindex$foldindex
missing = OCVdata
Y_data = missing[1,,]

out = softImpute(Y_data, rank.max = 40,lambda = 80)
length(out$d)
out = softImpute(Y_data, rank.max = 40,lambda = 30)
length(out$d)
out = softImpute(Y_data, rank.max = 40,lambda = 20)
length(out$d)
out = softImpute(Y_data, rank.max = 40,lambda = 10)
length(out$d)
out = softImpute(Y_data, rank.max = 40,lambda = 9)
length(out$d)
```

so it seems 10 or 11 should be a good choice which "Ideally lambda should be chosen so that the solution reached has rank slightly less than rank.max"

## OCV choice

for the grids we choose in our paper, we use 0-100 with ten grids

```{r}
library(softImpute)
N = dim(Y_data)[1]
P = dim(Y_data)[2]
ngrids = 10
fold = 10
K = 40
c_s = seq(0,100,len=ngrids)
cvout = CVPMD_softImpute(Y_data,c_s,K ,fold , method = "softImpute")
cvout
```

the result fit the authors' suggestion well (we choose $\lambda = 11$).

But can we try smaller grids with OCV

```{r}
library(softImpute)
N = dim(Y_data)[1]
P = dim(Y_data)[2]
ngrids = 10
fold = 10
K = 40
c_s = seq(0,15,len=ngrids)
cvout = CVPMD_softImpute(Y_data,c_s,K ,fold , method = "softImpute")
cvout
```

At this time, we can see that we choose $\lambda = 5$ this time.

## check the result

```{r}
out = softImpute(Y_data, rank.max = 40,lambda = 10)
miss_hat = out$u %*% diag(out$d) %*% t(out$v)
SSE = 0
for(j in 1:fold){
      SSE = SSE + sum((Y[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]] -
                               miss_hat[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]])^2,na.rm = TRUE)
}
sqrt(SSE/sum(is.na(Y_data)))
```

```{r}
out = softImpute(Y_data, rank.max = 40,lambda = 5)
miss_hat = out$u %*% diag(out$d) %*% t(out$v)
SSE = 0
for(j in 1:fold){
      SSE = SSE + sum((Y[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]] -
                               miss_hat[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]])^2,na.rm = TRUE)
}
sqrt(SSE/sum(is.na(Y_data)))
```

we can see that the result is slightly improved.
